Apache Flink\texttrademark: Unified Stream and Batch Processing \\ in a Single Engine

### 2015-11-27 batch.tex
small fix

\subsection{Batch Iterations on top of Dataflows}
\label{sec:batch-iterations}
Most batch dataflow approaches execute iterations by submitting a new job for each iteration or by adding additional nodes to a running DAG \cite{DBLP:journals/pvldb/BuHBE10, DBLP:conf/hotcloud/ZahariaCFSS10}. Flink keeps the DAG-based runtime and scheduler, and allows for special ``head'' and ``tail'' tasks to signify the beginning and end of each iteration, that are connected with feedback edges. Flink's iterations are competitive with specialized graph engines \cite{DBLP:journals/pvldb/EwenTKM12}. This is achieved through the concept of  \emph{delta} iterations, which exploit sparse computational dependencies, and are used as the basis for Gelly, Flink's Graph API. Finally, Flink offers Bulk Synchronous Parallel (BSP) as well as Stale Synchronous Parallel (SSP) iterations in its DataSet API.

>>>>

\subsection{Batch Iterations on top of Dataflows}
\label{sec:batch-iterations}
Most batch dataflow approaches execute iterations by submitting a new job for each iteration or by adding additional nodes to a running DAG \cite{DBLP:journals/pvldb/BuHBE10, DBLP:conf/hotcloud/ZahariaCFSS10}. Flink keeps the DAG-based runtime and scheduler, and allows for special ``head'' and ``tail'' tasks that are connected with feedback edges, to signify the beginning and end of each iteration. Flink's iterations are competitive with specialized graph engines \cite{DBLP:journals/pvldb/EwenTKM12}. This is achieved through the concept of  \emph{delta} iterations, which exploit sparse computational dependencies, and are used as the basis for Gelly, Flink's Graph API. Finally, Flink offers Bulk Synchronous Parallel (BSP) as well as Stale Synchronous Parallel (SSP) iterations in its DataSet API.

### 2015-11-27 intro.tex
small fix

Apache Flink follows a paradigm that embraces data stream processing as the unifying model for real-time analysis, continuous streams, and batch processing, both in the programming model and in the execution engine. In combination with durable message queues that allow quasi arbitrary replay of data streams (like Apache Kafka\footnote{http://kafka.apache.org/} or Amazon Kinesis\footnote{https://aws.amazon.com/kinesis/}), stream processing programs make no difference between processing the latest events in real-time, continuously aggregating data periodically in large windows, or processing terabytes of historical data: these types of computations simply start their processing at different points in the durable stream, and maintain different forms of state during the computation. Through a highly flexible windowing mechanism, Flink programs can compute both early and approximate, as well as delayed and accurate results in the same operation, obviating the need to combine different systems for the two use cases. Flink supports different notions of time (event time, ingestion time, processing time) in order to give programmer high flexibility in defining how events should be correlated.

At the same time, Flink acknowledges that there is, and will be, a need for dedicated batch processing (dealing with static data sets). Complex queries over static data are still a good match for a batch processing abstraction. Furthermore, batch processing is still needed both for legacy implementations of streaming use cases, and for analysis applications where no efficient algorithms are yet known that perform this kind of processing on streaming data. Batch programs are special cases of streaming programs, where the stream is finite, and order/time of records does not matter (all records implicitly belong to one all-encompassing window). However, to support batch use cases with a competitive ease and performance, Flink has a specialized API for processing static data sets, uses specialized data structures and algorithms for the batch versions of operators like join or grouping, and uses dedicated scheduling strategies. The result is that Flink presents itself as a full-fledged and efficient batch processor on top of a streaming runtime, including libraries for graph analysis and machine learning. 
Originating from the Stratosphere project \cite{stratosphere}, Flink is a top-level project of the Apache Software Foundation that is developed and supported by a large and lively community (of more than 140 open source contributors at the time of this writing\footnote{For an update list, refer to \url{https://github.com/apache/flink/graphs/contributors}}), and is used in production in several companies.

>>>>

Apache Flink follows a paradigm that embraces data stream processing as the unifying model for real-time analysis, continuous streams, and batch processing, both in the programming model and in the execution engine. In combination with durable message queues that allow quasi arbitrary replay of data streams (like Apache Kafka or Amazon Kinesis), stream processing programs make no difference between processing the latest events in real-time, continuously aggregating data periodically in large windows, or processing terabytes of historical data: these types of computations simply start their processing at different points in the durable stream, and maintain different forms of state during the computation. Through a highly flexible windowing mechanism, Flink programs can compute both early and approximate, as well as delayed and accurate results in the same operation, obviating the need to combine different systems for the two use cases. Flink supports different notions of time (event time, ingestion time, processing time) in order to give programmer high flexibility in defining how events should be correlated.

At the same time, Flink acknowledges that there is, and will be, a need for dedicated batch processing (dealing with static data sets). Complex queries over static data are still a good match for a batch processing abstraction. Furthermore, batch processing is still needed both for legacy implementations of streaming use cases, and for analysis applications where no efficient algorithms are yet known that perform this kind of processing on streaming data. Batch programs are special cases of streaming programs, where the stream is finite, and order/time of records does not matter (all records implicitly belong to one all-encompassing window). However, to support batch use cases with a competitive ease and performance, Flink has a specialized API for processing static data sets, uses specialized data structures and algorithms for the batch versions of operators like join or grouping, and uses dedicated scheduling strategies. The result is that Flink presents itself as a full-fledged and efficient batch processor on top of a streaming runtime, including libraries for graph analysis and machine learning. 
Originating from the Stratosphere project \cite{stratosphere}, Flink is a top-level project of the Apache Software Foundation that is developed and supported by a large and lively community (of more than 140 open source contributors at the time of this writing, and is used in production in several companies.

### 2015-11-27 streaming.tex
small fix

Note that streams are already partitioned on a key before windowing, so the above is a \textit{local operator} that does not require coordination between machines. This mechanism can be used to implement a wide variety of windowing functionality \cite{akidau2015dataflow}.

\subsection{Asynchronous Streaming Iterations}
Stream loops are crucial for several applications. Common examples are feedback loops for incrementally building and training machine learning models, reinforcement learning, and executing graph algorithms. Flink supports loops through explicit feedback edges in its dataflow graph, allowing to implement asynchronous iterative/incremental algorihtms. In Flink, feedback streams have special semantics and are treated in a way that is transparent to its scheduler: feedback streams do not act as dependencies in the dataflow graph, allowing Flink to employ simple DAG scheduling techniques while the contents of feedback streams are treated as operator state, in order to simplify fault tolerance \cite{carbone2015lightweight}. Flink allows constructing loops on top of its DataStream API, and can trivially subsume structured loops such as differential dataflows \cite{murray2013naiad}. \autoref{sec:batch-iterations} describes iterations support for batch analytics.

>>>>

Note that if the stream above is partitioned on a key before  windowing, the window operation above is \textit{local} and thus, it does not require coordination between machines. This mechanism can be used to implement a wide variety of windowing functionality \cite{akidau2015dataflow}.

\subsection{Asynchronous Streaming Iterations}
Stream loops are crucial for several applications. Common examples are feedback loops for incrementally building and training machine learning models, reinforcement learning, and executing graph algorithms. Flink supports loops through explicit feedback edges in its dataflow graph, allowing the implementation of asynchronous iterative/incremental algorihtms. In Flink, feedback streams have special semantics and are treated in a way that is transparent to its scheduler: feedback streams do not act as dependencies in the dataflow graph, allowing Flink to employ simple DAG scheduling techniques while the contents of feedback streams are treated as operator state, in order to simplify fault tolerance \cite{carbone2015lightweight}. Flink allows constructing loops on top of its DataStream API, and can trivially subsume structured loops such as differential dataflows \cite{murray2013naiad}. \autoref{sec:batch-iterations} describes iterations support for batch analytics.
